---
title: "Exploratory Analysis"
output:
  html_document:
    fig_caption: yes
    css: md-modest.css
  word_document:
    toc: yes
    fig_caption: yes
  pdf_document:
    
    toc: yes
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
library(lattice)
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

\

# The Lattice ^[plotting]{.smallcaps}^ system

Basic call be like: `xyplot(y ~ x | f * g, data)`, where f and g are "conditioning" factors or groups.

```{r}
library(datasets)

xyplot(Ozone ~ Wind, data = airquality)
```

Lattice plot according to category---better if it's a factor, I guess.

```{r}

airquality <- transform(airquality, Month = factor(Month))

xyplot(Ozone ~ Wind | Month, data = airquality, layout = c(5, 1))
```

# ggplot2

### `qplot()` (or, quick plot)

The basic grammar of plots is that they are made of *aeshtetics* (size, kinds of points, colors) and *geoms* (points, lines...). `ggplot()` is the actual, generic function.

```{r}
library(ggplot2)

qplot(displ, hwy, data = mpg)

# then you can change the color:
qplot(displ, hwy, data = mpg, color = drv)

# Add geoms, one at a time. When setting arg explicitly, you need to specify
# each and every one of them, including those that would have been default.
qplot(displ, hwy, data = mpg, geom = c("point", "smooth"))
```

### Histos

Single variable → histogram.

```{r}

qplot(hwy, data = mpg, fill = drv)
```

### Facets

Panel plots in facets using `facets` argument; variables on LHS are rows, RHS columns. Indicate "no var" with `.`.

```{r}

qplot(displ, hwy, data = mpg, facets = . ~ drv)


qplot(hwy, data = mpg, facets = drv ~ .)

```

## Real `ggplot2`

Component of a god-fearing upstanding ggplot:

-   A **data frame**

-   **Aesthetic** mappings: map data to color, shape, size, etc...

-   `geoms`: objects like points, lines, shapes

-   **Facets** for conditional plots

-   **`stats`**: statistical transformations like bins, quantiles, regression...

-   **`scales`**: the scale that aesthetics use, when applicable

-   **Coordinates**
